---
title: "a4_task2_Rammy_Salem"
author: "Rammy Salem"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# load packages
library(tidyverse)
library(janitor)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(effsize)
```

```{r}
# read csv file
snowshoe_hares <- read_csv(here("data","bonanza_hares.csv"))
```

# **Introduction**
#### **[add intro last]**

# **Data and analyses**
#### This data set contains the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 in the Bonanza Creek Experimental Forest located approximately 20 km southwest of Fairbanks, Alaska. The data were collected through the Bonanza Creek Long-Term Ecological Research program (Kielland, Chapin, Ruess, & Bonanza Creek LTER, 2017). **[provide an overview of the main analysis, statistical methods and tools used.]**

# **Exploratory Findings**
### **Annual juvenile hare trap counts**
```{r}
juvenile_snowshoe_hares <- snowshoe_hares %>% # Start from snowshoe hares data
  filter(age == "j") %>%  # Filtered to only include observations for juvenile snowshoe hares
  mutate(grid_full = case_when( # Create a new column called site_full that contains the full name for each of the five sites
    grid == "bonbs" ~ "Bonanza Black Spruce",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian",
  ))
```


```{r}
juvenile_snowshoe_hares_traps <- juvenile_snowshoe_hares %>% # Start from juvenile snowshoe hares data
  mutate(date = mdy(date)) %>% # Convert the date to class “Date”
  mutate(year = year(date)) %>% # Create a new column that only contains the year
  group_by(year) %>% # Group by site and year
  count(year) %>% # Count observations by group
  as.data.frame(juvenile_snowshoe_hares_traps) # Convert juvenile snowshoe hares list to data frame
```

```{r, include = FALSE}
 juvenile_snowshoe_hares_traps_table <- juvenile_snowshoe_hares_traps %>%
  summarize(mean_trappings = mean(n),# Find mean size
            median_trappings = median(n), # Find median size
            sd_trappings = sd(n), # Find standard deviation size
            n = sum(n)) # Find sample size
```

```{r, fig.cap= "Figure 1. Total number of juvenile snowshoe hare trappings by year"}
ggplot(data = juvenile_snowshoe_hares_traps,aes(x = year, y = n)) + # select x-axis variable
  geom_bar(stat="identity") + #create bar graph
  scale_x_continuous(name="Year", # Add x-axis label
                     breaks = seq(1998,2012,1), # specify breaks
                     limits=c(1998, 2012)) + # specify limits
  scale_y_continuous(name = "Number of Juvenile Snowshoe Hares", # Add y-axis labels
                     breaks = seq(0,130,10), # specify breaks)
                     limits=c(0, 130)) # specify limits
```
#### As shown in Figure 1, juvenile snowshoe hare trappings generally decreased from 1998-2012. The largest number of trappings occurred in 1999 (*n* = 126), and no juvenile snowshoe hare trappings were recorded for the years 1998, 2002, 2009, and 2012. A total of 378 juvenile snowshoe hares were trapped during this period, and the average number of snowshoe hare trappings per year was 31.5. Since the number of trappings are impacted by effort such as the number of days and number of traps set per year, one recommendation for the researchers would be to determine and agree upon a standardized number of days and trappings per year in order to control for the effects of these exogenous variables. 

### **Visualize juvenile hare weights**
```{r}
juvenile_snowshoe_hares_weights <- juvenile_snowshoe_hares %>% # Start from juvenile snowshoe hares data
  group_by(grid, sex) %>% # group by site and by sex
  summarize(mean_weight = mean(weight, na.rm = T),# Find mean weight
            median_weight = median(weight, na.rm = T), # Find median weight
            sd_weight = sd(weight, na.rm = T)) # Find standard deviation weight
  
```
```{r}
juvenile_snowshoe_hares_weights_by_sex <- juvenile_snowshoe_hares %>% # Start from juvenile snowshoe hares data
  group_by(sex) %>% # group by site and by sex
  summarize(mean_weight = mean(weight, na.rm = T),# Find mean weight
            median_weight = median(weight, na.rm = T), # Find median weight
            sd_weight = sd(weight, na.rm = T)) # Find standard deviation weight
```
```{r}
juvenile_snowshoe_hares_weights_by_site <- juvenile_snowshoe_hares %>% # Start from juvenile snowshoe hares data
  group_by(grid) %>% # group by site and by sex
  summarize(mean_weight = mean(weight, na.rm = T),# Find mean weight
            median_weight = median(weight, na.rm = T), # Find median weight
            sd_weight = sd(weight, na.rm = T)) # Find standard deviation weight
```


```{r,  fig.cap = "Figure 2. Distribution of weights for juvenile snowshoe hares by sex and site. *Note*. NA = sex could not be confirmed."}
mean_disp <- mean(juvenile_snowshoe_hares$weight, na.rm = T)
# mean_disp

ggplot(data = juvenile_snowshoe_hares, aes(x = sex, y = weight, color = sex)) + # select x- and y-axis variables
  geom_boxplot() + #create boxplot
  facet_wrap(~grid_full) + # facet wrapped by site
  scale_x_discrete(labels = c('Female', 'Male', 'NA')) +
  scale_y_continuous(name = "Juvenile Snowshoe Hare Weight (grams)", # Add y-axis label
                     breaks = seq(0,1600,200), # specify breaks)
                     limits=c(0, 1600)) + # specify limits 
  geom_hline(yintercept = mean_disp, linetype = "dashed") +
  labs(title = "Juvenile Snowshoe Hare Weights by Sex and Site", # Add graph title
       x = "Sex", # Added x-axis label
       color = "Sex")
```

#### As shown in Figure 2, juvenile snowshoe hare weights in grams were, on average, higher for males (*M* = 945.86, *SD* = 333.22) than females (*M* = 855.39, *SD* = 292.25) and hares whose sex could not be identified (*M* = 614.55, *SD* = 357.59) across all three trapping sites. In addition, juvenile snowshoe hares trapped in the Bonanza Black Spruce site weighed more (*M* = 1036.17, *SD* = 273.93), on average, than juvenile snowshoe hares trapped in the Bonanza Mature (*M* = 878.05, *SD* = 338.92) and Bonanza Riparian (*M* = 862.59, *SD* = 313.15) sites.

### **Juvenile weight comparison (male & female snowshoe hares).**
```{r}
juvenile_snowshoe_hares_weights_table_by_sex <- juvenile_snowshoe_hares %>% # Start from juvenile snowshoe hares data set
  drop_na("sex") %>% # remove the NAs in sex
  group_by(sex) %>% # group by sex
  summarize(mean_weight = mean(weight, na.rm = T),# Find mean weight
            median_weight = median(weight, na.rm = T), # Find median weight
            sd_weight = sd(weight, na.rm = T), # Find sample size
            sample_size = n()) # Find standard deviation weight

juvenile_snowshoe_hares_weights_table_by_sex %>% 
  kable(col.names = c("Sex", # Rename columns
                      "Mean weight (g)",
                      "Median weight (g)",
                      "SD weight (g)",
                      "Sample size (n)"),
        caption = "Table 1. Descriptive Statistics for Weights of Male and Female Snowshoe Hares") %>%  # Add table caption
  kable_styling()  # Use kableExtra to make table look nice
```
```{r, include = FALSE}
# Faceted Histograms
ggplot(data = juvenile_snowshoe_hares, aes(x = weight)) + 
  geom_histogram() + 
  facet_wrap(~sex)
```

```{r, include = FALSE}
# Faceted QQ Plots
ggplot(data = juvenile_snowshoe_hares, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```


```{r, include = FALSE}
# Create vectors containing observations for each sample (females and males)
female_sample <- juvenile_snowshoe_hares %>% 
  filter(sex == "f") %>% # filter for only juvenile female snowshoe hares
  pull(weight) # create vector of size data

male_sample <- juvenile_snowshoe_hares %>%
  filter(sex == "m") %>% # filter only control treatment
  pull(weight) # Create vector of size data

juvenile_snowshoe_hares_ttest <- t.test(x = male_sample, female_sample) # Perform a two-sample t-test to compare means between the two groups
juvenile_snowshoe_hares_ttest # display results

juvenile_snowshoe_hares_tidytable <- broom::tidy(juvenile_snowshoe_hares_ttest) # get results using broom package
juvenile_snowshoe_hares_tidytable # display results

cohen.d(male_sample, female_sample, na.rm = T)
```

#### Results of an independent samples *t*-test comparing juvenile snowshoe hare weights in grams of males vs. females revealed that males (*M* = 945.86, *SD* = 333.22) weighed significantly more than females (*M* = 855.39, *SD* = 292.25), *t*(325.02) = 2.71, *p* = .007, *d* = .29. The effect size for this analysis (*d* = .29) was found to exceed Cohen’s (1988) convention for a small effect (*d* = .20). 


### **Relationship between juvenile weight & hind foot length**

```{r, fig.cap = "Figure 3. Relationship between Juvenile Snoeshoe Hare Weight & Hind Foot Length"}
ggplot(data = juvenile_snowshoe_hares, aes(x = hindft, # select x-axis variable
                                         y = weight)) + # Select y-axis variable
  geom_point() + # create scatterplot
  scale_y_continuous(name = "Hare Weight (grams)", # Add y-axis label
                     breaks = seq(0,1600,200), # specify breaks
                     limits=c(0, 1600)) + # specify limits
  scale_x_continuous(name = "Hare Hindfoot Length (mm)", # Add x-axis label
                     breaks = seq(0,160,10), # specify breaks
                     limits=c(0, 160)) # specify limits
```

```{r, include = FALSE}
snowshoe_hare_lm <- lm(weight ~ hindft, data = juvenile_snowshoe_hares)
summary(snowshoe_hare_lm)
```

```{r, include = FALSE}
#get the model outputs in nice data frame format
snowshoe_hare_lm_tidy <- broom::tidy(snowshoe_hare_lm)
snowshoe_hare_lm_tidy 
```

```{r, include = FALSE}
# Get the intercept: 
snowshoe_hare_int <- snowshoe_hare_lm_tidy$estimate[1]
snowshoe_hare_int
```

```{r, include = FALSE}
# Then to get the flipper_length coefficient:
snowshoe_hare_coef <- snowshoe_hare_lm_tidy$estimate[2]
snowshoe_hare_coef
```

```{r, include = FALSE}
# Metrics at a glance: 
snowshoe_hare_lm_out <- broom::glance(snowshoe_hare_lm)
snowshoe_hare_lm_out
```


```{r, fig.cap = "Figure 4. Simple Linear Regression for Relationship between Juvenile Snoeshoe Hare Weight & Hind Foot Length"}
ggplot(data = juvenile_snowshoe_hares, aes(x = hindft, # select x-axis variable
                                         y = weight)) + # Select y-axis variable
  geom_point() + # create scatterplot
  scale_y_continuous(name = "Hare Weight (grams)", # Add y-axis label
                     breaks = seq(0,1600,200), # specify breaks
                     limits=c(0, 1600)) + # specify limits
  scale_x_continuous(name = "Hare Hindfoot Length (mm)", # Add x-axis label
                     breaks = seq(0,160,10), # specify breaks
                     limits=c(0, 160)) + # specify limits
  geom_smooth(method = "lm") + # add a linear model to the scatterplot
  ggpubr::stat_regline_equation(label.x = 70, label.y = 1150) # add regression equation
```

```{r, include = FALSE}
# Explore model assumptions
plot(snowshoe_hare_lm)
```

```{r, include = FALSE}
snowshoe_hare_cor <- cor.test(juvenile_snowshoe_hares$hindft, juvenile_snowshoe_hares$weight)
snowshoe_hare_cor
```

#### Simple linear regression was used to explore the relationship between juvenile snowshoe hare hindfoot length (mm) and weight (g) including both male and female hares. A significant regression model was found ($\beta$ = `r round(snowshoe_hare_coef,3)`, *F*(`r snowshoe_hare_lm_out$df`,`r snowshoe_hare_lm_out$df.residual`) = `r round(snowshoe_hare_lm_out$statistic,1)`, p < .001) with an R^2^ of `r round(snowshoe_hare_lm_out$r.squared,3)`. Pearson's correlation indicated a moderate positive correlation between juvenile snowshoe hare hindfoot length and weight (*r* = .55, t(247) = 10.26, p < 0.001).As for model assumptions, the variance of the residuals appeared relatively constant across the entire span of the model, suggesting homoscedasticity, but the QQ-plot revealed that the distribution was trending away from linearity, so this may be a cause for concern. 



#### **Summary**
Juvenile snowshoe hare hindfoot length predicts hare weight such that, on average, the longer the hindfoot, the larger the weight. On average, we expect weight to increase by 9.52 grams for each 1 mm increase in hindfoot length.

#### **Citation**: Kielland, K., F.S. Chapin, R.W. Ruess, & Bonanza Creek LTER. (2017). *Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present* ver 22. Environmental Data Initiative. (https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14)